### 2018-03-01 eval.Rmd
new env vars, biomark ratio

---
title: "Pteropod biomarker analysis"
author: ""
output: 
  html_document:
    code_folding: hide
self_contained: yes
---

```{r, message = F, warning = F}
library(knitr)
opts_chunk$set(warning = FALSE, message = FALSE, fig.path = 'figs/', dev.args = list(bg = 'transparent'), eval = T)

library(tidyverse)
library(vegan)
library(ggord)

source("R/funcs.R")

data(envdat)
data(biodat)

envchr <- c('Ara', 'O2', 'pH', 'Fluor', 'CO3', 'pCO2', 'Sal', 'TA', 'Temp', 'Latitude')
biochr <- c('CAT', 'GR', 'GSHonGSSG', 'GST', 'LPX', 'ORAC', 'SOD', 'ORACvLPX')
```

Pairs plot of biomarkers and environmental data
```{r, fig.height = 8, fig.width = 8}
dat <- biodat %>% 
  left_join(envdat, by = 'CTD') %>% 
  mutate_if(is.character, as.numeric) %>% 
  gather('var', 'val', -CTD) %>% 
  group_by(CTD, var) %>% 
  summarise(val = mean(val, na.rm = TRUE)) %>% 
  ungroup %>% 
  spread(var, val)

pairs(dat[, -1], gap = 0, cex = 1, col = 'grey')
```

RDA triplot of biomarkers and environmental data
```{r, fig.height = 4, fig.width = 5}
dat_frm <- dat %>% 
  mutate(
    CAT = ifelse(is.nan(CAT), mean(CAT, na.rm = TRUE), CAT)
  ) %>% 
  data.frame %>% 
  column_to_rownames('CTD') %>%
  decostand(method = 'range')

env <- select(dat_frm, one_of(envchr))
bio <- select(dat_frm, one_of(biochr))

mod <- rda(bio, env)
ggord(mod, ptslab = T, parse = T)
```

```{r}

# pearson, spearman correlation between biomarkers and env variables
cr <- dat_frm %>%
  gather('env_lab', 'env_val', one_of(envchr)) %>% 
  gather('bio_lab', 'bio_val', one_of(biochr)) %>% 
  group_by(env_lab, bio_lab) %>% 
  summarise(
    sp_cr = round(cor.test(env_val, bio_val, method = 'spearman')$estimate, 2),
    sp_pv = p_ast(cor.test(env_val, bio_val, method = 'spearman')$p.value),
    pr_cr = round(cor.test(env_val, bio_val, method = 'pearson')$estimate, 2),
    pr_pv = p_ast(cor.test(env_val, bio_val, method = 'pearson')$p.value)
  ) %>% 
  unite('sp_cr', sp_cr, sp_pv, sep = ' ') %>% 
  unite('pr_cr', pr_cr, pr_pv, sep = ' ')

sp_cr <- cr %>% 
  select(-pr_cr) %>% 
  spread(env_lab, sp_cr)

pr_cr <- cr %>% 
  select(-sp_cr) %>% 
  spread(env_lab, pr_cr)
```

# {.tabset}

## Spearman correlations
```{r}
knitr::kable(sp_cr, caption = 'Spearman rank correlations')
```

## Pearson correlations
```{r}
knitr::kable(pr_cr, caption = 'Pearson correlations')
```

# {.tabset}

Pairwise models with interactions for each biomarker
```{r}
# variables to model, chr strings only
env_cmb <- names(env) %>% 
  combn(2) %>% 
  t %>% 
  data.frame(stringsAsFactors = F) %>%
  crossing(names(bio), .) %>% 
  t %>% 
  data.frame(stringsAsFactors = F) %>% 
  as.list

# models
mods <- env_cmb %>%
  map(function(x){

    # formula
    frm <- paste(x[1], '~', x[2], '*', x[3]) %>% 
      as.formula

    # model, formatted
    mod <- lm(frm, data = dat_frm) %>% 
      summary %>% 
      .$coefficients %>% 
      .[-1, c(1, 4)] %>% 
      data.frame %>% 
      rownames_to_column('env_lab') %>% 
      mutate(
        bio_lab = x[1],
        Est = round(Estimate, 2), 
        Pvl = p_ast(Pr...t..)
      ) %>% 
      select(bio_lab, env_lab, Est, Pvl)

    return(mod)

  }) %>% 
  enframe('Model') %>% 
  unnest %>% 
  mutate(Model = gsub('^X', 'mod', Model)) %>% 
  group_by(bio_lab) %>% 
  nest
```

## CAT
```{r}
knitr::kable(filter(mods, bio_lab %in% 'CAT')$data[[1]], caption = 'CAT')
```

## GR
```{r}
knitr::kable(filter(mods, bio_lab %in% 'GR')$data[[1]], caption = 'GR')
```

## GSHonGSSG
```{r}
knitr::kable(filter(mods, bio_lab %in% 'GSHonGSSG')$data[[1]], caption = 'GSHonGSSG')
```

## GST
```{r}
knitr::kable(filter(mods, bio_lab %in% 'GST')$data[[1]], caption = 'GST')
```

## LPX
```{r}
knitr::kable(filter(mods, bio_lab %in% 'LPX')$data[[1]], caption = 'LPX')
```

## ORAC
```{r}
knitr::kable(filter(mods, bio_lab %in% 'ORAC')$data[[1]], caption = 'ORAC')
```

## SOD
```{r}
knitr::kable(filter(mods, bio_lab %in% 'SOD')$data[[1]], caption = 'SOD')
```

## ORACvLPX
```{r}
knitr::kable(filter(mods, bio_lab %in% 'ORACvLPX')$data[[1]], caption = 'ORACvLPX')
```