### 2018-03-02 index.Rmd
removed figs folder for rmarkdown shiny server

---
title: "Pteropod biomarker analysis"
author: ""
output: 
  html_document:
    code_folding: hide
self_contained: yes
runtime: shiny
---

```{r, message = F, warning = F}
library(knitr)
opts_chunk$set(warning = FALSE, message = FALSE, dev.args = list(bg = 'transparent'), eval = T)

library(tidyverse)
library(vegan)
library(ggord)
library(scales)
library(effects)
library(shiny)

source("R/funcs.R")

data(envdat)
data(biodat)

envchr <- c('pCO2', 'pH', 'CO3', 'Ara', 'O2', 'Temp')
biochr <- c('CAT', 'GR', 'GSHonGSSG', 'GST', 'LPX', 'ORAC', 'SOD', 'ORACvLPX')
```

Pairs plot of biomarkers and environmental data
```{r, fig.height = 8, fig.width = 8}
dat <- biodat %>% 
  left_join(envdat, by = 'CTD') %>% 
  mutate_if(is.character, as.numeric) %>% 
  select(one_of(c(envchr, biochr, 'CTD'))) %>% 
  gather('var', 'val', -CTD) %>% 
  group_by(CTD, var) %>% 
  summarise(val = mean(val, na.rm = TRUE)) %>%
  ungroup %>% 
  spread(var, val)

pairs(dat[, -1], gap = 0, cex = 1, col = 'grey')
```

RDA triplot of biomarkers and environmental data
```{r, fig.height = 4, fig.width = 5, eval = T}
dat_frm <- dat %>% 
  mutate(
    CAT = ifelse(is.nan(CAT), mean(CAT, na.rm = TRUE), CAT)
  ) %>% 
  data.frame %>% 
  column_to_rownames('CTD')

env <- select(dat_frm, one_of(envchr)) %>% 
  decostand(method = 'range')
bio <- select(dat_frm, one_of(biochr)) %>% 
  decostand(method = 'range')

mod <- rda(bio, env)
ggord(mod, ptslab = T, parse = T)
```

```{r}

# pearson, spearman correlation between biomarkers and env variables
cr <- dat_frm %>%
  gather('env_lab', 'env_val', one_of(envchr)) %>% 
  gather('bio_lab', 'bio_val', one_of(biochr)) %>% 
  group_by(env_lab, bio_lab) %>% 
  summarise(
    sp_cr = round(cor.test(env_val, bio_val, method = 'spearman')$estimate, 2),
    sp_pv = p_ast(cor.test(env_val, bio_val, method = 'spearman')$p.value),
    pr_cr = round(cor.test(env_val, bio_val, method = 'pearson')$estimate, 2),
    pr_pv = p_ast(cor.test(env_val, bio_val, method = 'pearson')$p.value)
  ) %>% 
  unite('sp_cr', sp_cr, sp_pv, sep = ' ') %>% 
  unite('pr_cr', pr_cr, pr_pv, sep = ' ')

sp_cr <- cr %>% 
  select(-pr_cr) %>% 
  spread(env_lab, sp_cr)

pr_cr <- cr %>% 
  select(-sp_cr) %>% 
  spread(env_lab, pr_cr)
```

# {.tabset}

## Spearman correlations
```{r}
knitr::kable(sp_cr, caption = 'Spearman rank correlations')
```

```{r spheat, fig.height = 4, fig.width = 5}
spplo <- cr %>% 
  select(-pr_cr) %>% 
  separate(sp_cr, c('cor', 'sig'), sep = ' ') %>% 
  mutate(
    cor = as.numeric(cor), 
    bio_lab = factor(bio_lab, levels = rev(unique(bio_lab)))
  )

pbase <- theme(
  panel.grid.major = element_blank(), 
  panel.grid.minor = element_blank(), 
  axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1, size = 8), 
  axis.text.y = element_text(size = 8),
  legend.position = 'top', 
  plot.margin = grid::unit(c(2, 2, 2, 2), 'pt'), 
  strip.background = element_blank(), 
  strip.text.y = element_text(angle = 0, hjust = 0, vjust = 0.5)
  )

p <- ggplot(spplo, aes(y = bio_lab, x = env_lab, fill = cor)) + 
  geom_tile(colour = 'black') + 
  pbase +
  scale_y_discrete('Biomarker', expand = c(0, 0)) + 
  scale_x_discrete('Environment', expand = c(0, 0)) +
  scale_fill_gradient2('Correlation', low = muted("green"), mid = "white", high = muted("blue"), midpoint = 0) +
  guides(fill = guide_colourbar(barheight = 0.5, barwidth = 5, label.theme = element_text(size = 6, angle = 0)))
p
```

## Pearson correlations
```{r}
knitr::kable(pr_cr, caption = 'Pearson correlations')
```

```{r prheat, fig.height = 4, fig.width = 5}
prplo <- cr %>% 
  select(-sp_cr) %>% 
  separate(pr_cr, c('cor', 'sig'), sep = ' ') %>% 
  mutate(
    cor = as.numeric(cor), 
    bio_lab = factor(bio_lab, levels = rev(unique(bio_lab)))
  )

pbase <- theme(
  panel.grid.major = element_blank(), 
  panel.grid.minor = element_blank(), 
  axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1, size = 8), 
  axis.text.y = element_text(size = 8),
  legend.position = 'top', 
  plot.margin = grid::unit(c(2, 2, 2, 2), 'pt'), 
  strip.background = element_blank(), 
  strip.text.y = element_text(angle = 0, hjust = 0, vjust = 0.5)
  )

p <- ggplot(prplo, aes(y = bio_lab, x = env_lab, fill = cor)) + 
  geom_tile(colour = 'black') + 
  pbase +
  scale_y_discrete('Biomarker', expand = c(0, 0)) + 
  scale_x_discrete('Environment', expand = c(0, 0)) +
  scale_fill_gradient2('Correlation', low = muted("green"), mid = "white", high = muted("blue"), midpoint = 0) +
  guides(fill = guide_colourbar(barheight = 0.5, barwidth = 5, label.theme = element_text(size = 6, angle = 0)))
p
```

# {.tabset}

Pairwise models with interactions for each biomarker
```{r}
# variables to model, chr strings only
env_cmb <- names(env) %>% 
  combn(2) %>% 
  t %>% 
  data.frame(stringsAsFactors = F) %>%
  crossing(names(bio), .) %>% 
  t %>% 
  data.frame(stringsAsFactors = F) %>% 
  as.list

# models
mod_all <- env_cmb %>%
  map(function(x){

    # formula
    frm <- paste(x[1], '~', x[2], '*', x[3]) %>% 
      as.formula

    # model only
    lm(frm, data = dat_frm)

  }) %>% 
  enframe('Model', 'Modobj') %>% 
  mutate(
    Model = gsub('^X', 'mod', Model), 
    data = map(Modobj, function(x){

      # response
      yvar <- all.vars(formula(x))[1]

      # rsq
      rsq <- summary(x) %>% 
        .$r.squared

      # coeff summary
      out <- x %>% 
        summary %>% 
        .$coefficients %>% 
        .[-1, c(1, 4)] %>% 
        data.frame %>% 
        rownames_to_column('env_lab') %>% 
        mutate(
          bio_lab = yvar,
          Rsq = round(rsq, 2),
          Est = round(Estimate, 2), 
          Pvl = p_ast(Pr...t..)
        ) %>% 
        select(bio_lab, env_lab, Rsq, Est, Pvl)

    return(out)

    })
  ) 
mods <- mod_all %>% 
  unnest(data)

```

```{r}
selectInput("mod", "Choose model:", paste0('mod', seq(1:nrow(mods))))

md <- reactive({

  md <- filter(mod_all, Model %in% input$mod) %>% 
  .$Modobj %>% 
  .[[1]]

  return(md)

})

levs1 <- reactive({

  levs1 <- quantile(md()$model[, 2],  c(0.25, 0.75)) %>% 
    list
  names(levs1) <- names(md()$model)[2]
  return(levs1)

})

levs2 <- reactive({

  levs2 <- quantile(md()$model[, 3], c(0.25, 0.75)) %>% 
    list
  names(levs2) <- names(md()$model)[3]
  return(levs2)

})

trm <- reactive({

  md() %>% coefficients %>% names %>% .[4]

})

renderPrint({summary(md())})

renderPlot({
  toplo <- md()$model %>% 
    gather('var', 'val', -1)

  ggplot(toplo, aes_string(x='val', y=names(toplo)[1])) +
    geom_point() + 
    facet_wrap(~var, ncol = 2, scales = 'free_x') + 
    theme_bw(base_family = 'serif') + 
    geom_smooth(method = 'lm') + 
    theme(axis.title.x = element_blank())

}, height = 300, width = 600)

renderPlot({
  plot(effect(trm(), md(), xlevels = levs1()), multiline = T)
}, height = 300, width = 600)

renderPlot({
  plot(effect(trm(), md(), xlevels = levs2()), multiline = T)
}, height = 300, width = 600)
```

## CAT
```{r}
knitr::kable(filter(mods, bio_lab %in% 'CAT'), caption = 'CAT')
```

## GR
```{r}
knitr::kable(filter(mods, bio_lab %in% 'GR'), caption = 'GR')
```

## GSHonGSSG
```{r}
knitr::kable(filter(mods, bio_lab %in% 'GSHonGSSG'), caption = 'GSHonGSSG')
```

## GST
```{r}
knitr::kable(filter(mods, bio_lab %in% 'GST'), caption = 'GST')
```

## LPX
```{r}
knitr::kable(filter(mods, bio_lab %in% 'LPX'), caption = 'LPX')
```

## ORAC
```{r}
knitr::kable(filter(mods, bio_lab %in% 'ORAC'), caption = 'ORAC')
```

## SOD
```{r}
knitr::kable(filter(mods, bio_lab %in% 'SOD'), caption = 'SOD')
```

## ORACvLPX
```{r}
knitr::kable(filter(mods, bio_lab %in% 'ORACvLPX'), caption = 'ORACvLPX')
```