### 2018-03-05 physio.Rmd

---
title: "Pteropod physiological analysis"
author: ""
output: 
  html_document:
    code_folding: hide
self_contained: yes
runtime: shiny
---

```{r, message = F, warning = F}
library(knitr)
opts_chunk$set(warning = FALSE, message = FALSE, dev.args = list(bg = 'transparent'), eval = T)

library(tidyverse)
library(vegan)
library(ggord)
library(scales)
library(effects)
library(shiny)
library(plotly)

source("R/funcs.R")

data(envdat)
data(ptedat)

envchr <- c('pCO2', 'pH', 'CO3', 'Ara', 'O2', 'Temp')
ptechr <- c('abu', 'dis', 'len')
```

Pairs plot of physiological parameters and environmental data
```{r, fig.height = 8, fig.width = 8}
dat <- ptedat %>% 
  select(one_of('CTD', ptechr)) %>% 
  left_join(envdat, by = 'CTD') %>% 
  select(one_of(c('CTD', envchr, ptechr)))

pairs(dat[, -1], gap = 0, cex = 1, col = 'grey')
```

Pairs plot of physiological parameters and biomarkers
```{r, fig.height = 8, fig.width = 8}
biopr <- ptedat %>% 
  select(-CTD, -Latitude)

pairs(biopr, gap = 0, cex = 1, col = 'grey')
```

RDA triplot of physiological parameters and environmental data
```{r, fig.height = 4, fig.width = 5, eval = T}
pca_frm <- dat %>% 
  na.omit %>% 
  data.frame %>% 
  column_to_rownames('CTD')

env <- select(pca_frm, one_of(envchr)) %>% 
  decostand(method = 'range')
pte <- select(pca_frm, one_of(ptechr)) %>% 
  decostand(method = 'range')

mod <- rda(pte, env)
ggord(mod, ptslab = T, parse = T)
```

```{r}
# pearson, spearman correlation between physiological parameters and env variables
cr <- dat %>%
  gather('env_lab', 'env_val', one_of(envchr)) %>% 
  gather('pte_lab', 'pte_val', one_of(ptechr)) %>% 
  group_by(env_lab, pte_lab) %>% 
  summarise(
    sp_cr = round(cor.test(env_val, pte_val, method = 'spearman')$estimate, 2),
    sp_pv = p_ast(cor.test(env_val, pte_val, method = 'spearman')$p.value),
    pr_cr = round(cor.test(env_val, pte_val, method = 'pearson')$estimate, 2),
    pr_pv = p_ast(cor.test(env_val, pte_val, method = 'pearson')$p.value)
  ) %>% 
  unite('sp_cr', sp_cr, sp_pv, sep = ' ') %>% 
  unite('pr_cr', pr_cr, pr_pv, sep = ' ')

sp_cr <- cr %>% 
  select(-pr_cr) %>% 
  spread(env_lab, sp_cr)

pr_cr <- cr %>% 
  select(-sp_cr) %>% 
  spread(env_lab, pr_cr)
```

# Correlation with environmental variables {.tabset}

## Spearman correlations, environmental variables
```{r}
knitr::kable(sp_cr, caption = 'Spearman rank correlations')
```

```{r, fig.height = 4, fig.width = 5}
spplo <- cr %>% 
  select(-pr_cr) %>% 
  separate(sp_cr, c('cor', 'sig'), sep = ' ') %>% 
  mutate(
    cor = as.numeric(cor), 
    pte_lab = factor(pte_lab, levels = rev(unique(pte_lab)))
  )

pbase <- theme(
  panel.grid.major = element_blank(), 
  panel.grid.minor = element_blank(), 
  axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1, size = 8), 
  axis.text.y = element_text(size = 8),
  legend.position = 'top', 
  plot.margin = grid::unit(c(2, 2, 2, 2), 'pt'), 
  strip.background = element_blank(), 
  strip.text.y = element_text(angle = 0, hjust = 0, vjust = 0.5)
  )

p <- ggplot(spplo, aes(y = pte_lab, x = env_lab, fill = cor)) + 
  geom_tile(colour = 'black') + 
  pbase +
  scale_y_discrete('Physio. param', expand = c(0, 0)) + 
  scale_x_discrete('Environment', expand = c(0, 0)) +
  scale_fill_gradient2('Correlation', low = muted("green"), mid = "white", high = muted("blue"), midpoint = 0) +
  guides(fill = guide_colourbar(barheight = 0.5, barwidth = 5, label.theme = element_text(size = 6, angle = 0)))
p
```

## Pearson correlations, environmental variables

```{r}
knitr::kable(pr_cr, caption = 'Pearson correlations')
```

```{r, fig.height = 4, fig.width = 5}
prplo <- cr %>% 
  select(-sp_cr) %>% 
  separate(pr_cr, c('cor', 'sig'), sep = ' ') %>% 
  mutate(
    cor = as.numeric(cor), 
    pte_lab = factor(pte_lab, levels = rev(unique(pte_lab)))
  )

p <- ggplot(prplo, aes(y = pte_lab, x = env_lab, fill = cor)) + 
  geom_tile(colour = 'black') + 
  pbase +
  scale_y_discrete('Physio. param', expand = c(0, 0)) + 
  scale_x_discrete('Environment', expand = c(0, 0)) +
  scale_fill_gradient2('Correlation', low = muted("green"), mid = "white", high = muted("blue"), midpoint = 0) +
  guides(fill = guide_colourbar(barheight = 0.5, barwidth = 5, label.theme = element_text(size = 6, angle = 0)))
p
```

# Correlation with biomarkers {.tabset}

```{r}
# pearson, spearman correlation between physiological parameters and biomarkers
cr <- ptedat %>% 
  gather('phy_lab', 'phy_val', one_of(ptechr)) %>% 
  gather('bio_lab', 'bio_val', -phy_lab, -phy_val, -CTD) %>% 
  group_by(phy_lab, bio_lab) %>% 
  summarise(
    sp_cr = round(cor.test(phy_val, bio_val, method = 'spearman')$estimate, 2),
    sp_pv = p_ast(cor.test(phy_val, bio_val, method = 'spearman')$p.value),
    pr_cr = round(cor.test(phy_val, bio_val, method = 'pearson')$estimate, 2),
    pr_pv = p_ast(cor.test(phy_val, bio_val, method = 'pearson')$p.value)
  ) %>% 
  unite('sp_cr', sp_cr, sp_pv, sep = ' ') %>% 
  unite('pr_cr', pr_cr, pr_pv, sep = ' ') %>% 
  ungroup

sp_cr <- cr %>% 
  select(-pr_cr) %>% 
  spread(bio_lab, sp_cr)

pr_cr <- cr %>% 
  select(-sp_cr) %>% 
  spread(bio_lab, pr_cr)
```

## Spearman correlations, biomarkers

```{r}
knitr::kable(sp_cr, caption = 'Spearman rank correlations')
```

```{r, fig.height = 4, fig.width = 5}
spplo <- cr %>% 
  select(-pr_cr) %>% 
  separate(sp_cr, c('cor', 'sig'), sep = ' ') %>% 
  mutate(
    cor = as.numeric(cor), 
    phy_lab = factor(phy_lab, levels = rev(unique(phy_lab)))
  )

p <- ggplot(spplo, aes(y = phy_lab, x = bio_lab, fill = cor)) + 
  geom_tile(colour = 'black') + 
  pbase +
  scale_y_discrete('Physio. param', expand = c(0, 0)) + 
  scale_x_discrete('Biomarker', expand = c(0, 0)) +
  scale_fill_gradient2('Correlation', low = muted("green"), mid = "white", high = muted("blue"), midpoint = 0) +
  guides(fill = guide_colourbar(barheight = 0.5, barwidth = 5, label.theme = element_text(size = 6, angle = 0)))
p
```

## Pearson correlations, biomarkers

```{r}
knitr::kable(pr_cr, caption = 'Pearson correlations')
```

```{r, fig.height = 4, fig.width = 5}
prplo <- cr %>% 
  select(-sp_cr) %>% 
  separate(pr_cr, c('cor', 'sig'), sep = ' ') %>% 
  mutate(
    cor = as.numeric(cor), 
    phy_lab = factor(phy_lab, levels = rev(unique(phy_lab)))
  )

p <- ggplot(prplo, aes(y = phy_lab, x = bio_lab, fill = cor)) + 
  geom_tile(colour = 'black') + 
  pbase +
  scale_y_discrete('Physio. param', expand = c(0, 0)) + 
  scale_x_discrete('Biomarker', expand = c(0, 0)) +
  scale_fill_gradient2('Correlation', low = muted("green"), mid = "white", high = muted("blue"), midpoint = 0) +
  guides(fill = guide_colourbar(barheight = 0.5, barwidth = 5, label.theme = element_text(size = 6, angle = 0)))
p
```

# {.tabset}

Pairwise models with interactions for each physiological parameter
```{r}
# variables to model, chr strings only
env_cmb <- names(env) %>% 
  combn(2) %>% 
  t %>% 
  data.frame(stringsAsFactors = F) %>%
  crossing(names(pte), .) %>% 
  t %>% 
  data.frame(stringsAsFactors = F) %>% 
  as.list

# models
mod_all <- env_cmb %>%
  map(function(x){

    # formula
    frm <- paste(x[1], '~', x[2], '*', x[3]) %>% 
      as.formula

    # model only
    lm(frm, data = dat)

  }) %>% 
  enframe('Model', 'Modobj') %>% 
  mutate(
    Model = gsub('^X', 'mod', Model), 
    data = map(Modobj, function(x){

      # response
      yvar <- all.vars(formula(x))[1]

      # rsq
      rsq <- summary(x) %>% 
        .$r.squared

      # coeff summary
      out <- x %>% 
        summary %>% 
        .$coefficients %>% 
        .[-1, c(1, 4)] %>% 
        data.frame %>% 
        rownames_to_column('env_lab') %>% 
        mutate(
          pte_lab = yvar,
          Rsq = round(rsq, 2),
          Est = round(Estimate, 2), 
          Pvl = p_ast(Pr...t..)
        ) %>% 
        select(pte_lab, env_lab, Rsq, Est, Pvl)

    return(out)

    })
  ) 
mods <- mod_all %>% 
  unnest(data)

```

```{r}
selectInput("mod", "Choose model:", paste0('mod', seq(1:nrow(mods))))

md <- reactive({

  md <- filter(mod_all, Model %in% input$mod) %>% 
  .$Modobj %>% 
  .[[1]]

  return(md)

})

levs1 <- reactive({

  levs1 <- quantile(md()$model[, 2],  c(0.25, 0.75)) %>% 
    list
  names(levs1) <- names(md()$model)[2]
  return(levs1)

})

levs2 <- reactive({

  levs2 <- quantile(md()$model[, 3], c(0.25, 0.75)) %>% 
    list
  names(levs2) <- names(md()$model)[3]
  return(levs2)

})

trm <- reactive({

  md() %>% coefficients %>% names %>% .[4]

})

prd_vl <- reactive({

  x <- md()$model %>% 
    .[, -1] %>% 
    data.frame %>% 
    as.list %>% 
    map(range) %>%
    map(function(x) seq(x[1], x[2], length = 100))

  nms <- names(x) 
  x <- crossing(x[[1]], x[[2]])
  names(x) <- nms

  prd_vl <- predict(md(), newdata = x) %>% 
    data.frame(., x)
  names(prd_vl)[1] <- all.vars(formula(md()))[1]

  return(prd_vl)

})

renderPrint({summary(md())})

renderPlot({
  toplo <- md()$model %>% 
    gather('var', 'val', -1)

  ggplot(toplo, aes_string(x='val', y=names(toplo)[1])) +
    geom_point() + 
    facet_wrap(~var, ncol = 2, scales = 'free_x') + 
    theme_bw(base_family = 'serif') + 
    geom_smooth(method = 'lm') + 
    theme(axis.title.x = element_blank())

}, height = 300, width = 600)

renderPlot({
  plot(effect(trm(), md(), xlevels = levs1()), multiline = T)
}, height = 300, width = 600)

renderPlot({
  plot(effect(trm(), md(), xlevels = levs2()), multiline = T)
}, height = 300, width = 600)

renderPlot({

  pbase <- theme(
    panel.grid.major = element_blank(), 
    panel.grid.minor = element_blank(), 
    axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1, size = 10), 
    axis.text.y = element_text(size = 10),
    legend.position = 'top', 
    plot.margin = grid::unit(c(2, 2, 2, 2), 'pt'), 
    strip.background = element_blank(), 
    strip.text.y = element_text(angle = 0, hjust = 0, vjust = 0.5)
  )

  p <- ggplot(prd_vl(), aes_string(y = names(prd_vl())[2], x = names(prd_vl())[3], fill = names(prd_vl())[1])) + 
    geom_tile(colour = 'black') + 
    theme_bw(base_family = 'serif') +
    pbase +
    scale_y_continuous(expand = c(0, 0)) +
    scale_x_continuous(expand = c(0, 0)) +
    scale_fill_distiller(palette = 'Spectral') +
    guides(fill = guide_colourbar(barheight = 0.5, barwidth = 15, label.theme = element_text(size = 12, angle = 0)))
  p
}, height = 700, width = 700)

renderPlotly({

  dat <- prd_vl() %>% 
    spread_(key_col = names(.)[2], value_col = names(.)[1]) %>% 
    .[,-1] %>% 
    as.matrix

  scene <- list(
    aspectmode = 'manual', 
    aspectratio = list(x = 1, y = 1, z = 0.5), 
    xaxis = list(title = names(prd_vl()[3])),
    yaxis = list(title = names(prd_vl()[2])),
    zaxis = list(title = names(prd_vl()[1]))
    )

  p <- plot_ly(z = ~dat) %>% add_surface(colors = rev(RColorBrewer::brewer.pal(11, 'Spectral'))) %>% 
    layout(scene = scene)
  p

})
```

## abundance
```{r}
knitr::kable(filter(mods, pte_lab %in% 'abu'), caption = 'abu')
```

## dissolution
```{r}
knitr::kable(filter(mods, pte_lab %in% 'dis'), caption = 'dis')
```

## length
```{r}
knitr::kable(filter(mods, pte_lab %in% 'len'), caption = 'len')
```