### 2018-03-10 expose.Rmd
added exposure analysis

---
title: "Pteropod exposure analysis"
author: ""
output: 
  html_document:
    code_folding: hide
    css: kable.css
self_contained: yes
runtime: shiny
---

```{r, message = F, warning = F}
library(knitr)
opts_chunk$set(warning = FALSE, message = FALSE, dev.args = list(bg = 'transparent'), eval = T)

library(tidyverse)
library(vegan)
library(ggord)
library(scales)
library(effects)
library(shiny)
library(plotly)

source("R/funcs.R")

data(envdat)
data(expdat)

envchr <- c('pCO2', 'pH', 'CO3', 'Ara', 'O2', 'Temp', 'Fluor')
```

A simple plot:
```{r, fig.height = 4, fig.width = 7}
toplo <- expdat %>% 
  unite('temp, pCO2', temp_trt, pco2_trt, sep = '/') %>% 
  mutate(
    `temp, pCO2` = factor(`temp, pCO2`, levels = c('lo/lo', 'lo/hi', 'hi/lo' ,'hi/hi'))
  )

ggplot(toplo, aes(x = `temp, pCO2`, y = alive)) +
  scale_y_continuous('% alive') +
  geom_boxplot(outlier.size = -1) +
  geom_jitter(alpha = 0.3, width = 0.15) + 
  theme_bw(base_family = 'serif')
```

Another simple plot:
```{r, fig.height = 4, fig.width = 7}
toplo <- expdat %>% 
  group_by(temp_trt, pco2_trt) %>% 
  summarize(
    ave = mean(alive, na.rm = T),
    lov = t.test(alive)$conf.int[1],
    hiv = t.test(alive)$conf.int[2]
  ) %>% 
  ungroup %>% 
  mutate(
    temp_trt = factor(temp_trt, levels = c('lo', 'hi')),
    pco2_trt = factor(pco2_trt, levels = c('lo', 'hi'))
  )

allv <- expdat %>% 
  mutate(
    temp_trt = factor(temp_trt, levels = c('lo', 'hi')),
    pco2_trt = factor(pco2_trt, levels = c('lo', 'hi'))
  ) %>%
  select(alive, temp_trt, pco2_trt, CTD) %>% 
  group_by(temp_trt, pco2_trt, CTD) %>%
  mutate(
    rep = 1, 
    rep = cumsum(rep)
    ) %>% 
  unite('uni', CTD, temp_trt, rep, remove = F)

ggplot(toplo, aes(x = pco2_trt, colour = temp_trt)) + 
  geom_line(data = allv, aes(y = alive, group = uni), linetype = 'dotted') +
  geom_point(size = 5, aes(y = ave)) + 
  geom_errorbar(aes(ymin = lov, ymax = hiv), width = 0.1, size = 1.5) +
  geom_line(aes(y = ave, group = temp_trt), size = 1.5) +
  scale_y_continuous('% alive') + 
  theme_bw(base_family = 'serif')
```

Run a simple test:
```{r}
mod <- lm(asin(alive / 100) ~ temp_trt * pco2_trt, data = expdat)
summary(mod)
```

No interaction:
```{r}
mod <- lm(asin(alive / 100) ~ temp_trt + pco2_trt, data = expdat)
summary(mod)
```

Compare with environmental parameters:
```{r}
expdat <- expdat %>% 
  left_join(envdat, by = 'CTD')
```

```{r}
selectInput("lns", "Choose a variable for line thickness:", envchr)

allv <- allv %>% 
  left_join(envdat, by = 'CTD')

plo <- reactive({

  names(allv)[names(allv) %in% input$lns] <- 'thick'

  p <- ggplot(toplo, aes(x = pco2_trt, colour = temp_trt)) + 
    geom_line(data = allv, aes(y = alive, group = uni, size = thick), alpha = 0.6) +
    geom_point(size = 5, aes(y = ave)) + 
    geom_errorbar(aes(ymin = lov, ymax = hiv), width = 0.1, size = 1.5) +
    geom_line(aes(y = ave, group = temp_trt), size = 1.5) +
    scale_y_continuous('% alive') + 
    theme_bw(base_family = 'serif') + 
    scale_size(input$lns, range = c(0.2, 1.5))

  return(p)

})

renderPlot({plo()}, height = 300, width = 600)

```

A model with environmental variable to model by treatment:
```{r}

selectInput("var", "Choose an environmental variable:", envchr)

# full
md <- reactive({

  frm <- formula(paste('asin(alive / 100) ~ (temp_trt + pco2_trt) * ', input$var))
  md <- lm(frm, data = allv)

  return(md)

})

# with select
md_sel <- reactive({

  frm <- formula(paste('asin(alive / 100) ~ (temp_trt + pco2_trt) * ', input$var))
  md_sel <- lm(frm, data = allv) %>% step(trace = 0)

  return(md_sel)

})

```

Full model:
```{r}
renderPrint({summary(md())})
```

Final selected model:
```{r}
renderPrint({summary(md_sel())})
```