### 2018-03-15 expose.Rmd
comps by treatment for exposure

---
title: "Pteropod exposure analysis"
author: ""
output: 
  html_document:
    code_folding: hide
    css: kable.css
self_contained: yes
runtime: shiny
---

```{r, message = F, warning = F}
library(knitr)
opts_chunk$set(warning = FALSE, message = FALSE, dev.args = list(bg = 'transparent'), eval = T)

library(tidyverse)
library(vegan)
library(ggord)
library(scales)
library(effects)
library(shiny)
library(plotly)

source("R/funcs.R")

data(envdat)
data(expdat)

envchr <- c('pCO2', 'pH', 'CO3', 'Ara', 'O2', 'Temp', 'Fluor')
```

A simple plot:
```{r, fig.height = 4, fig.width = 7}
toplo <- expdat %>% 
  unite('temp, pCO2', temp_trt, pco2_trt, sep = '/') %>% 
  mutate(
    `temp, pCO2` = factor(`temp, pCO2`, levels = c('lo/lo', 'lo/hi', 'hi/lo' ,'hi/hi'))
  )

ggplot(toplo, aes(x = `temp, pCO2`, y = alive)) +
  scale_y_continuous('% alive') +
  geom_boxplot(outlier.size = -1) +
  geom_jitter(alpha = 0.3, width = 0.15) + 
  theme_bw(base_family = 'serif')
```

Another simple plot:
```{r, fig.height = 4, fig.width = 7}
toplo <- expdat %>% 
  group_by(temp_trt, pco2_trt) %>% 
  summarize(
    ave = mean(alive, na.rm = T),
    lov = t.test(alive)$conf.int[1],
    hiv = t.test(alive)$conf.int[2]
  ) %>% 
  ungroup %>% 
  mutate(
    temp_trt = factor(temp_trt, levels = c('lo', 'hi')),
    pco2_trt = factor(pco2_trt, levels = c('lo', 'hi'))
  )

allv <- expdat %>% 
  mutate(
    temp_trt = factor(temp_trt, levels = c('lo', 'hi')),
    pco2_trt = factor(pco2_trt, levels = c('lo', 'hi'))
  ) %>%
  select(alive, temp_trt, pco2_trt, CTD) %>% 
  group_by(temp_trt, pco2_trt, CTD) %>%
  mutate(
    rep = 1, 
    rep = cumsum(rep)
    ) %>% 
  unite('uni', CTD, temp_trt, rep, remove = F)

ggplot(toplo, aes(x = pco2_trt, colour = temp_trt)) + 
  geom_line(data = allv, aes(y = alive, group = uni), linetype = 'dotted') +
  geom_point(size = 5, aes(y = ave)) + 
  geom_errorbar(aes(ymin = lov, ymax = hiv), width = 0.1, size = 1.5) +
  geom_line(aes(y = ave, group = temp_trt), size = 1.5) +
  scale_y_continuous('% alive') + 
  theme_bw(base_family = 'serif')
```

Run a simple test:
```{r}
mod <- lm(asin(alive / 100) ~ temp_trt * pco2_trt, data = expdat)
summary(mod)
```

No interaction:
```{r}
mod <- lm(asin(alive / 100) ~ temp_trt + pco2_trt, data = expdat)
summary(mod)
```

Compare with environmental parameters:
```{r}

toplo2 <- expdat %>% 
  left_join(envdat, by = 'CTD') %>% 
  unite(`temp, pCO2`, temp_trt, pco2_trt, sep = '/') %>% 
  mutate(
    `temp, pCO2` = factor(`temp, pCO2`, levels = c('lo/lo', 'lo/hi', 'hi/lo' ,'hi/hi')),
    alive_tr = asin(alive / 100)
  )

selectInput("vr", "Choose a variable to plot again survivorship:", envchr)
selectInput("trt", "Choose a treatment (temp, pCO2):", as.character(levels(toplo2$`temp, pCO2`)))

plovrid <- reactive({

  # assign variable selection to name
  names(toplo2)[names(toplo2) %in% input$vr] <- 'xval'

  # filter by treatment
  filt_toplo2 <- toplo2 %>% 
    filter(`temp, pCO2` %in% input$trt)

  p <- ggplot(filt_toplo2, aes(x = xval, y = alive)) + 
    geom_point(size = 5) + 
    scale_y_continuous('% alive') + 
    coord_cartesian(ylim = c(0, 105)) +
    theme_bw(base_family = 'serif') + 
    scale_x_continuous(input$vr) + 
    stat_smooth(method = 'lm') +
    ggtitle('identity')

  return(p)

})

plovr <- reactive({

  # assign variable selection to name
  names(toplo2)[names(toplo2) %in% input$vr] <- 'xval'

  # filter by treatment
  filt_toplo2 <- toplo2 %>% 
    filter(`temp, pCO2` %in% input$trt)

  p <- ggplot(filt_toplo2, aes(x = xval, y = alive_tr)) + 
    geom_point(size = 5) + 
    scale_y_continuous('% alive (arcsine)') + 
    coord_cartesian(ylim = c(0, 1.4)) +
    theme_bw(base_family = 'serif') + 
    scale_x_continuous(input$vr) + 
    stat_smooth(method = 'lm') +
    ggtitle('transformed')

  return(p)

})

renderPlot({plovrid()}, height = 300, width = 600)
renderPlot({plovr()}, height = 300, width = 600)
```

```{r, eval = T}
# all correlations
crs <- toplo2 %>% 
  select(-CTD) %>% 
  filter(!`temp, pCO2` %in% 'hi/hi') %>% 
  gather('var', 'val', -`temp, pCO2`, -alive_tr, -alive) %>% 
  gather('alvvr', 'alvvl', alive_tr, alive) %>% 
  group_by(`temp, pCO2`, alvvr, var) %>% 
  nest %>% 
  mutate(

    crs = map(data, function(x){

      # variables
      vr1 <- x$alvvl
      vr2 <- x$val

      # pearson
      pr_ts <- cor.test(vr1, vr2, method = 'pearson')
      pr_cr <- round(pr_ts$estimate, 2)
      pr_pv <- p_ast(pr_ts$p.value)
      pr <- paste(pr_cr, pr_pv)

      # spearman
      sp_ts <- suppressWarnings(cor.test(vr1, vr2, method = 'spearman'))
      sp_cr <- round(sp_ts$estimate, 2)
      sp_pv <- p_ast(sp_ts$p.value)
      sp <- paste(sp_cr, sp_pv)

      out <- data.frame(pr = pr, sp = sp, stringsAsFactors = F)
      return(out)

    })
  ) %>% 
  select(-data) %>% 
  unnest %>% 
  gather('cortyp', 'corval', pr, sp) %>% 
  spread(var, corval) %>% 
  rename(response = alvvr) %>% 
  mutate(response = factor(response, levels = c('alive', 'alive_tr'), labels = c('identity', 'arcsine'))) %>% 
  arrange(cortyp, response, `temp, pCO2`)
```

```{r, eval = T}
knitr::kable(crs, caption = 'Pearson and Spearman correlations of survival (identity and arcsine) by variable')
```