### 2018-04-02 physio.Rmd
effects plots with confidence intervals, added observed data to surface plots

---
title: "Pteropod physiological analysis"
author: ""
output: 
  html_document:
    code_folding: hide
    css: kable.css
self_contained: yes
runtime: shiny
---

```{r, message = F, warning = F}
library(knitr)
opts_chunk$set(warning = FALSE, message = FALSE, dev.args = list(bg = 'transparent'), eval = T)

library(tidyverse)
library(vegan)
library(ggord)
library(scales)
library(effects)
library(shiny)
library(plotly)

source("R/funcs.R")

data(envdat)
data(ptedat)
data(phymod)

envchr <- c('pCO2', 'pH', 'CO3', 'Ara', 'O2', 'Temp', 'Fluor')
ptechr <- c('abu', 'dis', 'len', 'ty2', 'ty3', 'scr')
```

Pairs plot
```{r, fig.height = 8, fig.width = 8}
dat <- ptedat %>% 
  select(one_of('CTD', ptechr)) %>% 
  left_join(envdat, by = 'CTD') %>% 
  select(one_of(c('CTD', envchr, ptechr)))

pairs(dat[, -1], gap = 0, cex = 1, col = 'grey')
```

RDA triplot of biomarkers and environmental data
```{r, fig.height = 4, fig.width = 5, eval = T}
dat_frm <- dat %>% 
  data.frame %>% 
  remove_rownames %>% 
  column_to_rownames('CTD')

env <- dat_frm %>% 
  na.omit %>% 
  select(., one_of(envchr)) %>% 
  decostand(method = 'range')
pte <- dat_frm %>% 
  na.omit %>% 
  select(one_of(ptechr)) %>% 
  decostand(method = 'range')

mod <- rda(pte, env)
ggord(mod, ptslab = T, parse = T)
```

# Physiological correlations with biomarkers {.tabset}

```{r}
tochk <- ptedat %>% 
  select(-CTD, -Latitude)

# all correlations
crs <- crossing(var1 = names(tochk), var2 = names(tochk)) %>% 
  filter(var1 != var2) %>% 
  rownames_to_column() %>% 
  group_by(rowname) %>% 
  nest %>% 
  mutate(
    crs = map(data, function(x){

      # variables
      vr1 <- tochk[[x$var1]]
      vr2 <- tochk[[x$var2]]

      # pearson
      pr_ts <- cor.test(vr1, vr2, method = 'pearson')
      pr_cr <- round(pr_ts$estimate, 2)
      pr_pv <- p_ast(pr_ts$p.value)
      pr <- paste(pr_cr, pr_pv)

      # spearman
      sp_ts <- suppressWarnings(cor.test(vr1, vr2, method = 'spearman'))
      sp_cr <- round(sp_ts$estimate, 2)
      sp_pv <- p_ast(sp_ts$p.value)
      sp <- paste(sp_cr, sp_pv)

      out <- data.frame(pr = pr, sp = sp, stringsAsFactors = F)
      return(out)

    })
  ) %>% 
  unnest %>% 
  select(-rowname)

sp_cr <- crs %>% 
  select(var1, var2, sp) %>% 
  spread(var2, sp) %>% 
  data.frame(stringsAsFactors = F) %>% 
  remove_rownames %>% 
  column_to_rownames('var1')

pr_cr <- crs %>% 
  select(var1, var2, pr) %>% 
  spread(var2, pr) %>% 
  data.frame(stringsAsFactors = F) %>% 
  remove_rownames %>% 
  column_to_rownames('var1') 
```

## Spearman correlations

<div class="verysmall">

```{r}
knitr::kable(sp_cr, caption = 'Spearman rank correlations')
```

</div>

```{r , fig.height = 5, fig.width = 10}
spplo <- crs %>% 
  select(-pr) %>% 
  separate(sp, c('cor', 'sig'), sep = ' ') %>% 
  mutate(
    cor = as.numeric(cor), 
    var1 = factor(var1, levels = rev(unique(.$var1)))
  )

pbase <- theme(
  panel.grid.major = element_blank(), 
  panel.grid.minor = element_blank(), 
  axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1, size = 8), 
  axis.text.y = element_text(size = 8),
  legend.position = 'top', 
  plot.margin = grid::unit(c(2, 2, 2, 2), 'pt'), 
  strip.background = element_blank(), 
  strip.text.y = element_text(angle = 0, hjust = 0, vjust = 0.5), 
  panel.background = element_rect(fill = 'black')
  )

p <- ggplot(spplo, aes(y = var1, x = var2, fill = cor)) + 
  geom_tile(colour = 'black') + 
  pbase +
  scale_y_discrete('', expand = c(0, 0)) + 
  scale_x_discrete('', expand = c(0, 0)) +
  scale_fill_gradient2('Correlation', low = muted("green"), mid = "white", high = muted("blue"), midpoint = 0) +
  guides(fill = guide_colourbar(barheight = 0.5, barwidth = 5, label.theme = element_text(size = 6, angle = 0)))
p
```

## Pearson correlations
<div class="verysmall">

```{r}
knitr::kable(pr_cr, caption = 'Pearson correlations')
```

</div>

```{r , fig.height = 5, fig.width = 10}
prplo <- crs %>% 
  select(-sp) %>% 
  separate(pr, c('cor', 'sig'), sep = ' ') %>% 
  mutate(
    cor = as.numeric(cor), 
    var1 = factor(var1, levels = rev(unique(.$var1)))
  )

pbase <- theme(
  panel.grid.major = element_blank(), 
  panel.grid.minor = element_blank(), 
  axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1, size = 8), 
  axis.text.y = element_text(size = 8),
  legend.position = 'top', 
  plot.margin = grid::unit(c(2, 2, 2, 2), 'pt'), 
  strip.background = element_blank(), 
  strip.text.y = element_text(angle = 0, hjust = 0, vjust = 0.5), 
  panel.background = element_rect(fill = 'black')
  )

p <- ggplot(prplo, aes(y = var1, x = var2, fill = cor)) + 
  geom_tile(colour = 'black') + 
  pbase +
  scale_y_discrete('', expand = c(0, 0)) + 
  scale_x_discrete('', expand = c(0, 0)) +
  scale_fill_gradient2('Correlation', low = muted("green"), mid = "white", high = muted("blue"), midpoint = 0) +
  guides(fill = guide_colourbar(barheight = 0.5, barwidth = 5, label.theme = element_text(size = 6, angle = 0)))
p
```

# Physiological correlations with environment {.tabset}

```{r}
# all correlations
crs <- crossing(var1 = names(dat_frm), var2 = names(dat_frm)) %>% 
  filter(var1 != var2) %>% 
  rownames_to_column() %>% 
  group_by(rowname) %>% 
  nest %>% 
  mutate(
    crs = map(data, function(x){

      # variables
      vr1 <- dat_frm[[x$var1]]
      vr2 <- dat_frm[[x$var2]]

      # pearson
      pr_ts <- cor.test(vr1, vr2, method = 'pearson')
      pr_cr <- round(pr_ts$estimate, 2)
      pr_pv <- p_ast(pr_ts$p.value)
      pr <- paste(pr_cr, pr_pv)

      # spearman
      sp_ts <- suppressWarnings(cor.test(vr1, vr2, method = 'spearman'))
      sp_cr <- round(sp_ts$estimate, 2)
      sp_pv <- p_ast(sp_ts$p.value)
      sp <- paste(sp_cr, sp_pv)

      out <- data.frame(pr = pr, sp = sp, stringsAsFactors = F)
      return(out)

    })
  ) %>% 
  unnest %>% 
  select(-rowname)

sp_cr <- crs %>% 
  select(var1, var2, sp) %>% 
  spread(var2, sp) %>% 
  data.frame(stringsAsFactors = F) %>% 
  remove_rownames %>% 
  column_to_rownames('var1')

pr_cr <- crs %>% 
  select(var1, var2, pr) %>% 
  spread(var2, pr) %>% 
  data.frame(stringsAsFactors = F) %>% 
  remove_rownames %>% 
  column_to_rownames('var1') 
```

## Spearman correlations
<div class="verysmall">

```{r}
knitr::kable(sp_cr, caption = 'Spearman rank correlations')
```

</div>

```{r spheat, fig.height = 5, fig.width = 10}
spplo <- crs %>% 
  select(-pr) %>% 
  separate(sp, c('cor', 'sig'), sep = ' ') %>% 
  mutate(
    cor = as.numeric(cor), 
    var1 = factor(var1, levels = rev(unique(.$var1)))
  )

pbase <- theme(
  panel.grid.major = element_blank(), 
  panel.grid.minor = element_blank(), 
  axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1, size = 8), 
  axis.text.y = element_text(size = 8),
  legend.position = 'top', 
  plot.margin = grid::unit(c(2, 2, 2, 2), 'pt'), 
  strip.background = element_blank(), 
  strip.text.y = element_text(angle = 0, hjust = 0, vjust = 0.5), 
  panel.background = element_rect(fill = 'black')
  )

p <- ggplot(spplo, aes(y = var1, x = var2, fill = cor)) + 
  geom_tile(colour = 'black') + 
  pbase +
  scale_y_discrete('', expand = c(0, 0)) + 
  scale_x_discrete('', expand = c(0, 0)) +
  scale_fill_gradient2('Correlation', low = muted("green"), mid = "white", high = muted("blue"), midpoint = 0) +
  guides(fill = guide_colourbar(barheight = 0.5, barwidth = 5, label.theme = element_text(size = 6, angle = 0)))
p
```

## Pearson correlations
<div class="verysmall">

```{r}
knitr::kable(pr_cr, caption = 'Pearson correlations')
```

</div>

```{r prheat, fig.height = 5, fig.width = 10}
prplo <- crs %>% 
  select(-sp) %>% 
  separate(pr, c('cor', 'sig'), sep = ' ') %>% 
  mutate(
    cor = as.numeric(cor), 
    var1 = factor(var1, levels = rev(unique(.$var1)))
  )

pbase <- theme(
  panel.grid.major = element_blank(), 
  panel.grid.minor = element_blank(), 
  axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1, size = 8), 
  axis.text.y = element_text(size = 8),
  legend.position = 'top', 
  plot.margin = grid::unit(c(2, 2, 2, 2), 'pt'), 
  strip.background = element_blank(), 
  strip.text.y = element_text(angle = 0, hjust = 0, vjust = 0.5), 
  panel.background = element_rect(fill = 'black')
  )

p <- ggplot(prplo, aes(y = var1, x = var2, fill = cor)) + 
  geom_tile(colour = 'black') + 
  pbase +
  scale_y_discrete('', expand = c(0, 0)) + 
  scale_x_discrete('', expand = c(0, 0)) +
  scale_fill_gradient2('Correlation', low = muted("green"), mid = "white", high = muted("blue"), midpoint = 0) +
  guides(fill = guide_colourbar(barheight = 0.5, barwidth = 5, label.theme = element_text(size = 6, angle = 0)))
p
```

# Pairwise VIF among environmental variables

```{r}
vifv <- crossing(var1 = envchr, var2 = envchr) %>% 
  filter(var1 != var2) %>% 
  rownames_to_column() %>% 
  group_by(rowname) %>% 
  nest %>% 
  mutate(
    vifv = map(data, function(x){

    vr1 <- x[['var1']]
    vr2 <- x[['var2']]
    frm <- formula(paste(vr1, '~', vr2))
    mod <- lm(frm, dat_frm)
    rsq <- mod %>% 
      summary %>% 
      .$r.squared
    vif <- 1 / (1 - rsq)

    return(vif)

    })
  ) %>% 
  unnest %>% 
  select(-rowname) %>% 
  mutate(vifv = round(vifv, 2)) %>% 
  spread(var2, vifv) %>% 
  data.frame(stringsAsFactors = F) %>%
  remove_rownames %>% 
  column_to_rownames('var1') 
```

<div class="verysmall">

```{r}
knitr::kable(vifv, caption = 'VIF')
```

</div>

# {.tabset}

Pairwise models with interactions for each biomarker

```{r, fig.height = 9}
# model data
mods <- phymod %>% 
  unnest(data)
mod_all <- phymod

selectInput("mod", "Choose model:", paste0('mod', seq(1:nrow(mods))))

md <- reactive({

  md <- filter(mod_all, Model %in% input$mod) %>% 
  .$Modobj %>% 
  .[[1]]

  return(md)

})

levs1 <- reactive({

  levs1 <- quantile(md()$model[, 2],  c(0, 1)) %>% 
    list
  names(levs1) <- names(md()$model)[2]
  return(levs1)

})

levs2 <- reactive({

  levs2 <- quantile(md()$model[, 3], c(0, 1)) %>% 
    list
  names(levs2) <- names(md()$model)[3]
  return(levs2)

})

trm <- reactive({

  md() %>% coefficients %>% names %>% .[4]

})

prd_vl <- reactive({

  x <- md()$model %>% 
    .[, -1] %>% 
    data.frame %>% 
    as.list %>% 
    map(range) %>%
    map(function(x) seq(x[1], x[2], length = 100))

  nms <- names(x) 
  x <- crossing(x[[1]], x[[2]])
  names(x) <- nms

  prd_vl <- predict(md(), newdata = x) %>% 
    data.frame(., x)
  names(prd_vl)[1] <- all.vars(formula(md()))[1]

  return(prd_vl)

})

renderPrint({summary(md())})

renderPlot({
  toplo <- md()$model %>% 
    gather('var', 'val', -1)

  ggplot(toplo, aes_string(x='val', y=names(toplo)[1])) +
    geom_point() + 
    facet_wrap(~var, ncol = 2, scales = 'free_x') + 
    theme_bw(base_family = 'serif') + 
    geom_smooth(method = 'lm') + 
    theme(axis.title.x = element_blank())

}, height = 300, width = 600)

renderPlot({
  plot(effect(trm(), md(), xlevels = levs1()), multiline = T)
}, height = 300, width = 600)

renderPlot({
  plot(effect(trm(), md(), xlevels = levs2()), multiline = T)
}, height = 300, width = 600)

renderPlot({

  pbase <- theme(
    panel.grid.major = element_blank(), 
    panel.grid.minor = element_blank(), 
    axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1, size = 10), 
    axis.text.y = element_text(size = 10),
    legend.position = 'top', 
    plot.margin = grid::unit(c(2, 2, 2, 2), 'pt'), 
    strip.background = element_blank(), 
    strip.text.y = element_text(angle = 0, hjust = 0, vjust = 0.5)
  )

  p <- ggplot(prd_vl(), aes_string(y = names(prd_vl())[2], x = names(prd_vl())[3], fill = names(prd_vl())[1])) + 
    geom_tile(colour = 'black') + 
    theme_bw(base_family = 'serif') +
    pbase +
    scale_y_continuous(expand = c(0, 0)) +
    scale_x_continuous(expand = c(0, 0)) +
    scale_fill_distiller(palette = 'Spectral') +
    guides(fill = guide_colourbar(barheight = 0.5, barwidth = 15, label.theme = element_text(size = 12, angle = 0)))
  p
}, height = 700, width = 700)

renderPlotly({

  # observed data
  toplo <- md()$model
  names(toplo) <- c('z', 'x', 'y')

  # surface data
  dat <- prd_vl() %>% 
    spread_(key_col = names(.)[2], value_col = names(.)[1]) %>% 
    as.matrix
  x <- colnames(dat)[-1] %>% as.numeric
  y <- dat[, 1]
  dat <- dat[, -1]

  scene <- list(
    aspectmode = 'manual', 
    aspectratio = list(x = 1, y = 1, z = 0.6), 
    xaxis = list(title = names(prd_vl()[2])),
    yaxis = list(title = names(prd_vl()[3])),
    zaxis = list(title = names(prd_vl()[1]))
  )

  p <- plot_ly() %>% 
    add_surface(z = ~dat, x = ~x, y = ~y, colors = rev(RColorBrewer::brewer.pal(11, 'Spectral'))) %>% 
  add_markers(data = toplo, x = ~x, y = ~y, z = ~z) %>% 
    layout(scene = scene, width = 800, height = 400)
  p
})
```

## abundance
```{r}
knitr::kable(filter(mods, pte_lab %in% 'abu'), caption = 'abu')
```

## dissolution
```{r}
knitr::kable(filter(mods, pte_lab %in% 'dis'), caption = 'dis')
```

## type 2 dissolution
```{r}
knitr::kable(filter(mods, pte_lab %in% 'ty2'), caption = 'ty2')
```

## type 3 dissolution
```{r}
knitr::kable(filter(mods, pte_lab %in% 'ty3'), caption = 'ty3')
```

## scarring
```{r}
knitr::kable(filter(mods, pte_lab %in% 'scr'), caption = 'scr')
```

## length
```{r}
knitr::kable(filter(mods, pte_lab %in% 'len'), caption = 'len')
```