### 2018-07-23 figures.Rmd
O2 back in cellular models, updated figures.Rmd

---
output:
  html_document:
    keep_md: yes
    self_contained: no
    code_folding: hide
    css: styles.css
---

<a href="https://github.com/fawda123/Pteropod_biomarker/" class="github-corner" aria-label="View source on Github"><svg width="80" height="80" viewBox="0 0 250 250" style="fill:#FD6C6C; color:#fff; position: absolute; top: 0; border: 0; left: 0; transform: scale(-1, 1);" aria-hidden="true"><path d="M0,0 L115,115 L130,115 L142,142 L250,250 L250,0 Z"></path><path d="M128.3,109.0 C113.8,99.7 119.0,89.6 119.0,89.6 C122.0,82.7 120.5,78.6 120.5,78.6 C119.2,72.0 123.4,76.3 123.4,76.3 C127.3,80.9 125.5,87.3 125.5,87.3 C122.9,97.6 130.6,101.9 134.4,103.2" fill="currentColor" style="transform-origin: 130px 106px;" class="octo-arm"></path><path d="M115.0,115.0 C114.9,115.1 118.7,116.5 119.8,115.4 L133.7,101.6 C136.9,99.2 139.9,98.4 142.2,98.6 C133.8,88.0 127.5,74.4 143.8,58.0 C148.5,53.4 154.0,51.2 159.7,51.0 C160.3,49.4 163.2,43.6 171.4,40.1 C171.4,40.1 176.1,42.5 178.8,56.2 C183.1,58.6 187.2,61.8 190.9,65.4 C194.5,69.0 197.7,73.2 200.1,77.6 C213.8,80.2 216.3,84.9 216.3,84.9 C212.7,93.1 206.9,96.0 205.4,96.6 C205.1,102.4 203.0,107.8 198.3,112.5 C181.9,128.9 168.3,122.5 157.7,114.1 C157.9,116.9 156.7,120.9 152.7,124.9 L141.0,136.5 C139.8,137.7 141.6,141.9 141.8,141.8 Z" fill="currentColor" class="octo-body"></path></svg></a><style>.github-corner:hover .octo-arm{animation:octocat-wave 560ms ease-in-out}@keyframes octocat-wave{0%,100%{transform:rotate(0)}20%,60%{transform:rotate(-25deg)}40%,80%{transform:rotate(10deg)}}@media (max-width:500px){.github-corner:hover .octo-arm{animation:none}.github-corner .octo-arm{animation:octocat-wave 560ms ease-in-out}}</style>

# Pteropod stressor interaction: figures

```{r echo = T, message = F, warning = F}
library(tidyverse)
library(vegan)
library(ggord)
library(scales)
library(effects)
library(knitr)
library(gridExtra)
library(grid)
library(stargazer)

source("R/funcs.R")

opts_chunk$set(fig.align = 'center', message = F, echo = T, cache = F, dev = 'png', dev.args = list(family = 'serif'), dpi = 400, fig.pos = '!h', warning = F, background = 'white', out.width = '100%',
               fig.process = function(x) {
                 x2 = sub('-\\d+([.][a-z]+)$', '\\1', x)
                 if (file.rename(x, x2)) x2 else x
               })

data(envdat)
data(ptedat)
data(phymod)
data(biomod)

envchr <- c('Lat', 'pCO2', 'pH', 'Ara', 'O2', 'Temp', 'Fluor')
biochr <- c('CAT', 'GR', 'GSHonGSSG', 'GST', 'LPX', 'ORAC', 'SOD', 'ORACvLPX')
phychr <- c('abu', 'dis', 'len', 'ty2', 'ty3', 'scr')
```

```{r rda_bio}
# biomarker rda model
dat_bio <- ptedat %>% 
  select(one_of('CTD', biochr)) %>% 
  na.omit %>% 
  inner_join(envdat, by = 'CTD') %>% 
  select(one_of(c('CTD', envchr, biochr))) %>% 
  data.frame %>% 
  remove_rownames %>% 
  column_to_rownames('CTD')

envphy <- select(dat_bio, one_of(envchr)) %>% 
  select(-Lat) %>% 
  decostand(method = 'range')
ptephy <- select(dat_bio, one_of(biochr)) %>% 
  decostand(method = 'range')

biorda<- rda(ptephy, envphy)

# physiology rda model
dat_phy <- ptedat %>% 
  select(one_of('CTD', phychr)) %>% 
  na.omit %>% 
  inner_join(envdat, by = 'CTD') %>% 
  select(one_of(c('CTD', envchr, phychr))) %>% 
  data.frame %>% 
  remove_rownames %>% 
  column_to_rownames('CTD') %>% 
  rename(
    abundance = abu, 
    dissolution = dis, 
    length = len, 
    scarring = scr, 
    typeII = ty2, 
    typeIII = ty3
  )

envphy <- select(dat_phy, one_of(envchr)) %>% 
  select(-Lat) %>% 
  decostand(method = 'range')
ptephy <- select(dat_phy, -one_of(envchr)) %>% 
  decostand(method = 'range')

phyrda <- rda(ptephy, envphy)

```

```{r rdaplo, fig.height = 4.5, fig.width = 9}
# globals
coord_fix <- F
repel <- F
addsize <- 3
alpha <- 0.8
arrow <- 0.3
cols <- 'lightblue'

scalbio <- range(dat_bio$Lat)
scalbio <- seq(scalbio[1], scalbio[2], length  = 5)
scalphy <- range(dat_phy$Lat)
scalphy <- seq(scalphy[1], scalphy[2], length  = 5)

p1a <- ggord(biorda, ptslab = T, parse = T, repel = repel, coord_fix = coord_fix, addsize = addsize, size = dat_bio$Lat, sizelab = 'Latitude', alpha = alpha, arrow = arrow, ellipse = F) + 
  ggtitle('(a) Cellular endpoints') + 
  scale_size(breaks = scalbio, labels = round(scalbio, 1), range = c(1, 8))
p2a <- ggord(phyrda, ptslab = T, parse = T, repel = repel, coord_fix = coord_fix, addsize = addsize, size = dat_phy$Lat, sizelab = 'Latitude', alpha = alpha, arrow = arrow, ellipse = F) + 
  ggtitle('(b) Physiological endpoints') + 
  scale_size(breaks = scalphy, labels = round(scalphy, 1), range = c(1, 8))

grid.arrange(p1a, p2a, ncol = 2)
```

Fig. 2 Results of redundancy analyses for environmental variables with (a) cellular and (b) physiological endpoints of pteropod response to OA stressors.  Points are site locations in multivariate space with the size proportional to latitude. Separate RDAs were created for cellular and physiological endpoints because not all data were available across all stations.

```{r}
# physiology rda model
dat_cor <- ptedat %>% 
  select(one_of('CTD', phychr, biochr)) %>% 
  # na.omit %>% 
  left_join(envdat, by = 'CTD') %>% 
  data.frame %>% 
  remove_rownames %>% 
  column_to_rownames('CTD')

# all correlations
crs <- crossing(var1 = names(dat_cor), var2 = names(dat_cor)) %>% 
  filter(var1 != var2) %>% 
  rownames_to_column() %>% 
  group_by(rowname) %>% 
  nest %>% 
  mutate(
    crs = map(data, function(x){

      # variables
      vr1 <- dat_cor[[x$var1]]
      vr2 <- dat_cor[[x$var2]]

      # pearson
      pr_ts <- cor.test(vr1, vr2, method = 'pearson')
      pr_cr <- round(pr_ts$estimate, 2)
      pr_pv <- p_ast(pr_ts$p.value)
      pr <- paste(pr_cr, pr_pv)

      out <- data.frame(pr = pr, stringsAsFactors = F)
      return(out)

    })
  ) %>% 
  unnest %>% 
  select(-rowname)
```

```{r corplo, fig.height = 6, fig.width = 7}
levs <- c(sort(envchr), sort(biochr), sort(phychr))
labs <- c('Omega[ar]', 'Fluor', 'Lat', 'O[2]', 'pCO[2]', 'pH', 'Temp', 'CAT', 'GR', 'GSHonGSSG', 'GST', ' LPX', 'ORAC', 'ORACvLPX', 'SOD', 'abundance', 'dissolution', 'length', 'scarring', 'typeII', 'typeIII')
prplo <- crs %>% 
  separate(pr, c('cor', 'sig'), sep = ' ') %>% 
  filter(var1 %in% levs & var2 %in% levs) %>%  
  mutate(
    cor = as.numeric(cor), 
    var1 = factor(var1, levels = rev(levs), labels = rev(labs)), 
    var2 = factor(var2, levels = rev(levs), labels = rev(labs)), 
    sig = gsub('ns', '', sig)
  )

pbase <- theme(
  panel.grid.major = element_blank(), 
  panel.grid.minor = element_blank(), 
  axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1, size = 8), 
  axis.text.y = element_text(size = 8),
  legend.position = c(0.5, 1.12),
  legend.direction = 'horizontal',
  plot.margin = unit(c(4,4,0,0), "lines"),
  strip.background = element_blank(), 
  strip.text.y = element_text(angle = 0, hjust = 0, vjust = 0.5), 
  panel.background = element_rect(fill = 'black')
  )

outlab <- data.frame(
  y = c(3.5, 10.5, 18.5), 
  lab = c('Population/\nPhysiology', 'Cellular', 'Environment')
)

p <- ggplot(prplo) + 
  geom_tile(aes(y = var1, x = var2, fill = cor), colour = 'black') + 
  geom_text(aes(y = var1, x = var2, label = sig)) +
  annotation_custom(grob = textGrob(label = outlab$lab[1], hjust = 0, gp = gpar(cex = 0.7)),
                    ymin = outlab$y[1], ymax = outlab$y[1], xmin = 22, xmax = 22) +
  annotation_custom(grob = textGrob(label = outlab$lab[2], hjust = 0, gp = gpar(cex = 0.7)),
                    ymin = outlab$y[2], ymax = outlab$y[2], xmin = 22, xmax = 22) +  
  annotation_custom(grob = textGrob(label = outlab$lab[3], hjust = 0, gp = gpar(cex = 0.7)),
                    ymin = outlab$y[3], ymax = outlab$y[3], xmin = 22, xmax = 22) +
  annotation_custom(grob = textGrob(label = outlab$lab[1], hjust = 0.5, gp = gpar(cex = 0.7)),
                    xmin = outlab$y[1], xmax = outlab$y[1], ymin = 22.5, ymax = 22.5) +
  annotation_custom(grob = textGrob(label = outlab$lab[2], hjust = 0.5, gp = gpar(cex = 0.7)),
                    xmin = outlab$y[2], xmax = outlab$y[2], ymin = 22.5, ymax = 22.5) +  
  annotation_custom(grob = textGrob(label = outlab$lab[3], hjust = 0.5, gp = gpar(cex = 0.7)),
                    xmin = outlab$y[3], xmax = outlab$y[3], ymin = 22.5, ymax = 22.5) +
  pbase +
  scale_y_discrete('', expand = c(0, 0), labels = parse(text = rev(labs))) + 
  scale_x_discrete('', expand = c(0, 0), labels = parse(text = rev(labs))) +
  scale_fill_gradient2('Correlation', low = muted("blue"), mid = "white", high = muted("red"), midpoint = 0) +
  guides(fill = guide_colourbar(barheight = 0.5, barwidth = 5, label.theme = element_text(size = 6, angle = 0))) +
  geom_hline(yintercept = 6.5, size = 1.5) +
  geom_hline(yintercept = 14.5, size = 1.5) +
  geom_vline(xintercept = 6.5, size = 1.5) +
  geom_vline(xintercept = 14.5, size = 1.5)

# Code to override clipping
gt <- ggplot_gtable(ggplot_build(p))
gt$layout$clip[gt$layout$name == "panel"] <- "off"
grid.draw(gt)
```
Fig. 3 Correlation matrix of environmental variables, cellular response endpoints, and population (abundance) and physiological response endpoints for pteropods.  Darker red values are strong positive correlations and darker purple values are strong negative correlations.  Significance values at alpha = 0.05 are shown by stars (p < 0.05 \*, p < 0.005 \*\*).

```{r effbio, fig.height = 8, fig.width = 6, eval = T}
toplo <- biomod %>% 
  filter(Model %in% c('mod3', 'mod4', 'mod24')) %>% 
  mutate(
    cvar = c('pCO2', 'Temp', 'Temp'),
    xvar = c('Fluorescence', 'Omega[ar]', 'Omega[ar]'), 
    yvar = c('LPX', 'LPX', 'SOD'), 
    legr = c('pCO[2]', 'Temp', 'Temp'),
    fct = c(0.5, 0.95, 0.95)
    ) %>% 
  mutate(
    pldat = pmap(list(Modobj, cvar, fct), function(Modobj, cvar, fct){

      get_pldat(Modobj, cvar, fct = fct)

    }),
    plos = pmap(list(pldat, xvar, legr), function(pldat, xvar, legr){

      pl1 <- pldat[[1]]
      pl2 <- pldat[[2]]

      # colors
      cols <- RColorBrewer::brewer.pal(9, 'RdBu')

      # names for aes_string
      nms <- names(pl1)

      p1 <- ggplot() +
        geom_ribbon(data = pl1, aes_string(x = nms[3], ymin = 'lo', ymax = 'hi', group = nms[4]), alpha = 0.5, fill = 'grey') +
        geom_line(data = pl1, aes_string(x = nms[3], y = nms[1], group = nms[4], colour = nms[4]), size = 1) + 
        geom_text(data = pl2, aes(x= x, y = y, label = lab), hjust = 0) +
        theme_bw() +
        xlab(parse(text = xvar)) +
        scale_colour_gradientn(parse(text = legr), colours = cols) +
        scale_y_continuous(nms[1])

      return(p1)

    })
  )

p1 <- toplo$plos[[1]] + ggtitle('(b)')
p2 <- toplo$plos[[2]] + ggtitle('(a)')
p3 <- toplo$plos[[3]] + ggtitle('(c)')

pA <- ggplot_gtable(ggplot_build(p1))
pB <- ggplot_gtable(ggplot_build(p2))
pC <- ggplot_gtable(ggplot_build(p3))

maxWidth = grid::unit.pmax(pA$widths[2:3], pB$widths[2:3], pC$widths[2:3])

pA$widths[2:3] <- maxWidth
pB$widths[2:3] <- maxWidth
pC$widths[2:3] <- maxWidth

grid.arrange(
 pB, pA, pC, ncol = 1 
)

```
Fig. 4 Examples of model interactions of co-occuring environmental variables on pteropod cellular responses. Plot (a) shows a negative additive effect between aragonite saturation state (Ωar) and temperature on cellular toxicity (lipid peroxidation; LPX); plot (b) shows negative additive effects of food availability and pCO2 on LPX; plot (c) shows positive additive effects of temperature and Ωar on antioxidative activity (sodium dismutase, SOD). Both y-axes are transformed to conform to model output. Covarying environmental variables were held constant at the minimum and maximum values in the observed data.

```{r effphy, fig.height = 5.33, fig.width = 6, eval = T}
toplo <- phymod %>% 
  filter(Model %in% c('mod2', 'mod9')) %>% 
  mutate(
    cvar = c('Temp', 'Temp'),
    xvar = c('Omega[ar]', 'Omega[ar]'), 
    yvar = c('Abundance', 'Dissolution'), 
    legr = c('Temp', 'Temp'),
    pos = c('left', 'right'), 
    fct = c(0.92, 1.04), 
    hjs = c(0, 1)
    ) %>% 
  mutate(
    pldat = pmap(list(Modobj, cvar, pos, fct), function(Modobj, cvar, pos, fct){

      get_pldat(Modobj, cvar, pos = pos, fct = fct)

    }),
    plos = pmap(list(pldat, xvar, yvar, legr, hjs), function(pldat, xvar, yvar, legr, hjs){

      pl1 <- pldat[[1]]
      pl2 <- pldat[[2]]

      # colors
      cols <- RColorBrewer::brewer.pal(9, 'RdBu')

      # names for aes_string
      nms <- names(pl1)

      p1 <- ggplot() +
        geom_ribbon(data = pl1, aes_string(x = nms[3], ymin = 'lo', ymax = 'hi', group = nms[4]), alpha = 0.5, fill = 'grey') +
        geom_line(data = pl1, aes_string(x = nms[3], y = nms[1], group = nms[4], colour = nms[4]), size = 1) + 
        geom_text(data = pl2, aes(x= x, y = y, label = lab), hjust = hjs) +
        theme_bw() +
        xlab(parse(text = xvar)) +
        scale_colour_gradientn(parse(text = legr), colours = cols) +
        scale_y_continuous(yvar)

      return(p1)

    })
  )

p1 <- toplo$plos[[1]] + ggtitle('(b)')
p2 <- toplo$plos[[2]] + ggtitle('(a)')

pA <- ggplot_gtable(ggplot_build(p1))
pB <- ggplot_gtable(ggplot_build(p2))

maxWidth = grid::unit.pmax(pA$widths[2:3], pB$widths[2:3])

pA$widths[2:3] <- maxWidth
pB$widths[2:3] <- maxWidth

grid.arrange(
 pB, pA, ncol = 1 
)

```
Fig. 5 An example of model interactions of co-occuring environmental variables with biomineralization (shell dissolution) and population (abundance) responses. The interaction effects between aragonite saturation state (Ωar) and temperature on shell dissolution, as well as the interaction between Ωar and oxygen is synergistic (a, b, respectively). All y-axes are transformed to conform to model output. Covarying environmental variables were held constant at the minimum and maximum values in the observed data.

```{r, eval = T, results = 'asis'}
biotab <- biomod %>% 
  unnest(data) %>% 
  group_by(Model) %>% 
  mutate(n = n()) %>% 
  filter(n > 1) %>% 
  dplyr::select(-n)
biotab <- filter(biomod, Model %in% unique(biotab$Model))

cap <- 'Table 1: Model results for pteropod cellular response to pairs of co-occurring environmental variables. The estimated joint effects of variables and their significance in each model are shown.  Parameter estimates are included with overall model statistics (R-squared, F statistic) at the bottom. Models with the same cellular response measure can be identified with the column names at the top.  * p < 0.1, ** p < 0.05, *** p < 0.01'

stargazer(biotab$Modobj, type = 'html', title = cap, digits = 2, keep.stat = c('adj.rsq'), intercept.top = T, intercept.bottom = F, dep.var.caption = 'Cellular response measures', report = 'vc*')
```

```{r, results = 'asis'}
phytab <- phymod %>% 
  unnest(data) %>% 
  group_by(Model) %>% 
  mutate(n = n()) %>% 
  filter(n > 1) %>% 
  dplyr::select(-n)
phytab <- filter(phymod, Model %in% unique(phytab$Model))

cap <- 'Table 2: Model results for physiological and abundance responses to pairs of co-occurring environmental variables. The estimated joint effects of variables and their significance in each model are shown.  Parameter estimates are included with overall model statistics (R-squared, F statistic) at the bottom. Models with the same response measure can be identified with the column names at the top. * p < 0.1, ** p < 0.05, *** p < 0.01'

stargazer(phytab$Modobj, type = 'html', title = cap, digits = 2, keep.stat = c('adj.rsq'), intercept.top = T, intercept.bottom = F, dep.var.caption = 'Physiological response measures', dep.var.labels = c('Abundance', 'Shell dissolution'), report = 'vc*')

```