### 2021-08-10 exposure.Rmd
exposure rmd

---
output:
  html_document:
    self_contained: no
    code_folding: hide
    css: styles.css
---

```{r setup, warning = F, message = F}
library(knitr)
opts_chunk$set(warning = FALSE, message = FALSE, dev.args = list(bg = 'transparent'), eval = T)

library(tidyverse)
library(patchwork)
library(here)

data(expdat)
```

```{r, results = 'hide'}
toplo <- expdat %>% 
  unite('temp, pCO2', temp_trt, pco2_trt, sep = '/') %>% 
  mutate(
    `temp, pCO2` = factor(`temp, pCO2`, levels = c('lo/lo', 'lo/hi', 'hi/lo' ,'hi/hi'))
  )

p1 <- ggplot(toplo, aes(x = `temp, pCO2`, y = alive)) +
  scale_y_continuous('% alive') +
  geom_boxplot(outlier.size = -1) +
  geom_jitter(alpha = 0.3, width = 0.15) + 
  theme_bw(base_family = 'serif', base_size = 14) + 
  labs(x = 'Temp, pCO2')

toplo <- expdat %>% 
  group_by(temp_trt, pco2_trt) %>% 
  summarize(
    ave = mean(alive, na.rm = T),
    lov = t.test(alive)$conf.int[1],
    hiv = t.test(alive)$conf.int[2]
  ) %>% 
  ungroup %>% 
  mutate(
    temp_trt = factor(temp_trt, levels = c('lo', 'hi')),
    pco2_trt = factor(pco2_trt, levels = c('lo', 'hi'))
  )

allv <- expdat %>% 
  mutate(
    temp_trt = factor(temp_trt, levels = c('lo', 'hi')),
    pco2_trt = factor(pco2_trt, levels = c('lo', 'hi'))
  ) %>%
  select(alive, temp_trt, pco2_trt, CTD) %>% 
  group_by(temp_trt, pco2_trt, CTD) %>%
  mutate(
    rep = 1, 
    rep = cumsum(rep)
  ) %>% 
  unite('uni', CTD, temp_trt, rep, remove = F)

p2 <- ggplot(toplo, aes(x = pco2_trt, colour = temp_trt)) + 
  # geom_line(data = allv, aes(y = alive, group = uni), linetype = 'dotted') +
  geom_point(size = 5, aes(y = ave)) + 
  geom_errorbar(aes(ymin = lov, ymax = hiv), width = 0.1, size = 1.5) +
  geom_line(aes(y = ave, group = temp_trt), size = 1.5) +
  theme_bw(base_family = 'serif', base_size = 14) + 
  labs(
    x = 'pCO2', 
    y = '% alive', 
    colour = 'Temp'
  ) +
  theme(legend.position = c(0.9, 0.2))

p <- p1 + p2 + plot_layout(ncol = 1)

png(here('raw/survive.png'), height = 6.5, width = 6.5, family = 'serif', res = 300, units = 'in')
print(p)
dev.off()
```

[download figure](raw/survive.png)

```{r, out.width = '80%'}
knitr::include_graphics(here('raw/survive.png'))
```

Run a simple test:
```{r}
mod <- lm(asin(alive / 100) ~ temp_trt * pco2_trt, data = expdat)
summary(mod)
anova(mod)
```

No interaction:
```{r}
mod <- lm(asin(alive / 100) ~ temp_trt + pco2_trt, data = expdat)
summary(mod)
anova(mod)
```